A hidden Active Galactic Nuclei population: the first radio luminosity functions constructed by physical process

### 2024-11-07 src/tex/ms.tex
[showyourwork] overleaf sync

% Title
\title[A hidden AGN population]{A hidden Active Galactic Nuclei population: the first radio luminosity functions by physical process}

@@

\begin{abstract}
Both star formation (SF) and Active Galactic Nuclei (AGN) play an important role in galaxy evolution. Quantifying their relative importance in a statistical way can be done using radio luminosity functions. Until now these have relied on galaxy classifications, where sources with a mixture of radio emission from SF and AGN are labelled as either a star-forming galaxy or an AGN. This can lead to the under- or over-estimation of the relevance of AGN. Brightness temperature methods were demonstrated to be effective at separating radio emission from AGN and SF at 144 MHz using sub-arcsecond imaging with the International LOFAR Telescope (ILT). We use the combination of sub-arcsecond and arcsecond resolution imaging of 7,497 sources in the Lockman Hole and ELAIS-N1 fields to identify AGN components in the high resolution images and subtract them from the total flux density, leaving flux density from SF only. We construct, for the first time, radio luminosity functions by physical process, either SF or AGN activity, revealing a hidden AGN population at \llof $<10^{24}$\wphz . This population is 1.56$\pm$0.06 more than expected for $0.5<z<2.0$ when comparing to RLFs by galaxy classification. The star forming population is less impacted, with only 0.90$\pm$0.02 of the expected star formation. The contribution of `hidden' AGN to the overall radio population, especially at \llof $<10^{24}$\wphz , can have significant implications for the cosmic star formation rate and kinetic luminosity densities.
\end{abstract}

>>>>

% Title
\title[A hidden AGN population]{A hidden Active Galactic Nuclei population: the first radio luminosity functions constructed by physical process}

@@

\begin{abstract}
Both star formation (SF) and Active Galactic Nuclei (AGN) play an important role in galaxy evolution. Statistically quantifying their relative importance can be done using radio luminosity functions. Until now these relied on galaxy classifications, where sources with a mixture of radio emission from SF and AGN are labelled as either a star-forming galaxy or an AGN. This can cause the misestimation of the relevance of AGN. Brightness temperature measurements at 144 MHz with the International LOFAR telescope can separate radio emission from AGN and SF. We use the combination of sub-arcsec and arcsec resolution imaging of 7,497 sources in the Lockman Hole and ELAIS-N1 fields to identify AGN components in the sub-arcsec resolution images and subtract them from the total flux density, leaving flux density from SF only. We construct, for the first time, radio luminosity functions by physical process, either SF or AGN activity, revealing a hidden AGN population at \llof$<10^{24}$\wphz . This population is 1.56$\pm$0.06 more than expected for $0.5<z<2.0$ when comparing to RLFs by galaxy classification. The star forming population has only 0.90$\pm$0.02 of the expected SF. These ‘hidden’ AGN can have significant implications for the cosmic star formation rate and kinetic luminosity densities.
\end{abstract}
